---
title: "Diarrhea Recall Error Check- Kenya"
author: "Andrew Mertens"
date: "December 13, 2016"
output: pdf_document
---


```{r, include=FALSE}
#---------------------------------------
# preamble
#---------------------------------------
rm(list=ls())
library(metafor)
library(washb)
library(dplyr)


#---------------------------------------
# Load the analysis dataset
#---------------------------------------
d <- read.csv("C:/Users/andre/Dropbox/WBK-primary-analysis/Data/Final/Andrew/washb-kenya-diar.csv")

#---------------------------------------
# Subset the Data to Follow-up data only
#---------------------------------------
table(d$studyyear)

#---------------------------------------
# Exclude:
# * siblings who were born after enrollment
# * siblings who were >36 mos at enrollment
# * children with missing outcome data
#---------------------------------------


#subset the diarrhea to children <36 mos at enrollment
#or target children
ad <-
  d %>%
  subset(., studyyear>0 & dcohort==1) %>%
  subset(., !is.na(.$diar7d))

dim(ad)
table(ad$diar7d, ad$tr,ad$studyyear)
table(ad$tr,ad$studyyear)

# re-order the tr factor so Control 
ad$tr <- factor(ad$tr,levels=c("Passive Control","Water","Sanitation","Handwashing","WSH","Nutrition","Nutrition + WSH","Control"))
```

###Examine overlap between primary outcomes definition of diarrhea, and a definition that constrains 3+ stool and watery stool to occur on the same day. 
(Only using 2 day recall because no information on the day the abnormal stool occurred between days 3-7.)  

Table of primary outcome diarrhea definition compared to stricter, alternative definition:  

```{r, echo=F, comment=""}
#table(ad$altdiar7d,ad$diar7d)
#table(factor(ad$altdiar2d,levels=c("No diar","Diar")),factor(ad$diar2d, levels=c("No diar","Diar")), dnn=c("Alt def","Primary outcome def"))
table(ad$altdiar2d,ad$diar2d, dnn=c("Alt def","Primary outcome def"))

```

There are 152 less 2-day recall diarrhea cases in the analysis population if the diarrhea definition is constrained to have 3+ stool and watery stool occurr on the same day (5.6% of cases used in the primary outcomes analysis).



###Create a variable for terminated diarrhea cases between 7d and 2d recall periods

```{r, tidy=T, comment=""}
ad$diart <- ifelse(ad$diar2d==0 & ad$diar7d==1,1,0)
table(ad$diar2d,ad$diart, dnn=c("2d diarrhea","Terminated diarrhea"))  
```

There were 532 terminated cases where diarrhea was reported over the last 7 days, but not within the last 2 days.

 
```{r, include=F, comment=""}
#---------------------------------------
# Current:Terminated ratio difference function
#---------------------------------------
ctratio <- function(dcurr,dterm,tr) {
  # dcurr: indicator of a current diarrhea case
  # dterm: indicator of a terminated diarrhea case
  # tr   : assigned treatment group
  #curr <- tapply(dcurr,tr,sum)
  #term <- tapply(dterm,tr,sum)
  curr <- table(dcurr, tr)[2,]  
  term <-table(dterm,tr)[2,] 
  ctratio <- curr/term
  return(ctratio)
}

```

###Examine the ratio between current and terminated diarrhea across treatment arms:
```{r, echo=F, comment=""}
ctratio(ad$diar2d,ad$diart,ad$tr)
```

Roughly 5/1 ratio of ongoing (or recent) diarrhea cases to terminated cases within the last week. 


###Cross-tabs of final observations in the analysis, by survey round and study arm
```{r, echo=F, comment=""}
table(ad$tr,ad$diar2d, ad$studyyear, dnn=c("","2-day diarrhea recall","Study Year"))
table(ad$tr,ad$diar7d, ad$studyyear, dnn=c("","7-day diarrhea recall","Study Year"))
table(ad$tr,ad$diart, ad$studyyear, dnn=c("","Terminated diarrhea recall","Study Year"))

```



###Results of bootstrap the Current diarrhea/Terminated diarrhea ratio by re-sampling randomization blocks with replacement:

```{r, echo=F, cache=T}


set.seed(1349175)
nreps <- 1000
bsamp <- matrix(sample(unique(ad$block),size=length(unique(ad$block))*nreps,replace=TRUE),ncol=nreps)
ctratios <- matrix(rep(NA,nreps*8),ncol=8)
for(i in 1:nreps) {
  bd <- merge(ad,data.frame(block=bsamp[,i]),by="block",all.x=FALSE)
  ctratios[i,] <- ctratio(bd$diar2d,bd$diart,bd$tr)
}

# compute differences between each arm and control, then the mean and percentile 95% CIs
muctratio <- apply(ctratios,2,mean)[c(8,1:7)]
ctdiff <- ctratios[,1:7]-ctratios[,8]
ctmeans <- apply(ctdiff,2,mean)
ct95ci  <- apply(ctdiff,2,function(x) quantile(x,probs=c(0.025,0.975)))

res <- cbind(muctratio,c(NA,ctmeans),t(cbind(c(NA,NA),ct95ci)))
rownames(res) <- levels(ad$tr)[c(8,1:7)]
colnames(res) <- c("CTratio","CTratio diff","Lower 95% CI","Upper 95% CI")
```


```{r,echo=F, comment=""}
res
```


No evidence for significant difference in diarrhea recall between study arms.
